New lower bounds for Schur and weak Schur numbers

### 2021-04-08 article/article.tex

@@

>>>>

\par
These new results come from Rowley's approach for Ramsey graphs which is relatively new. Therefore, it would not surprise us
if lower bounds are later improved using better templates. Moreover, studying specifically \(S_+(n)\) and \(WS_+(n)\) numbers might be
of interest as they are closely related to Schur and weak Schur numbers.

\par
The fifth section gives new insight on the method that was formerly used to achieve new lower bounds. The principle behind it
might have discarded many interesting partititions of the search space and lowered the highest value that can be reached within it.
However, algorithms based on randomness such as Monte-Carlo algorithms may appear very useful if used in a search space with more
potential. For example, such an algorithm could be used to find better templates and improve inequalities and lower bounds for high values
of \(n\). This could be the subject fo a future work. Nevertheless, it would not help for the smallest values.

@@

\par
We would first like to address warm thanks to Fred Rowley who kindly answered our many questions and with who it was a pleasure
to exchange. A great thank-you to Marijn Heule who also took the time to respond to our requests. Finally, last but not least, we would
like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised us during the whole research and writing.

### 2021-04-08 article/schur.tex

>>>>

\textsc{Proof :} We denote by \(f\) the projection of the equivalence relation induced by the \(SF\)-template of lenght \(p\) 
and \(g\) the one induced by the sum-free partition of \([\![1,q]\!]\). Each equivalence class is represented by a single integer, therefore :

\[ f : [\![1,p]\!] \longrightarrow [\![1,k+1]\!] \text{ and } \forall (x,y) \in [\![1,p]\!]^2, \left\{
\begin{array}{ll}
	f(x) = f(y) \leqslant k \\
	x + y > p
\end{array}
\right.
\Longrightarrow f(x+y-p) \neq f(x)
\]
with also the sum-free condition : \[ \forall (x,y) \in [\![1,p]\!]^2 \text{, } f(x) = f(y)
\Longrightarrow f(x+y) \neq f(x) \]

\[g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } g(x) = g(y)
\Longrightarrow g(x+y) \neq g(x)
\]

We now define \(h : [\![1,pq]\!] \longrightarrow [\![1,n+k]\!] \) as follows :
\( \forall x \in [\![1,pq]\!] \), we write \(x = \alpha p - u\) where \(\alpha \in [\![1,q]\!] \) and \( u \in [\![0,p-1]\!] \).
This decomposition if of course unique. If \(f(p-u) \leqslant k\), we set \(h(x) = f(p-u)\), else (i.e \(f(p-u) = k + 1) \) \(h(x) = k + g(\alpha) \). 
Prooving that the partition of \([\![1,pq]\!]\) induced by \(h\) is sum-free will complete the proof.

Let \(x,y \in [\![1,pq]\!]\), with \(h(x) = h(y)\) and \(x+y \leqslant pq\). We write \(x = \alpha p - u\) and \(y = \beta p - v\). \\
We simply need \(h(x+y) \neq h(x)\). \\

We first examine the case where \(h(x) \leqslant k\). We assume \(h(x+y) \leqslant k\), otherwise we are fine. 
Thus by definition of \(h\), we have \(h(x) = f(p-u) = f(p-v) = h(y)\).
\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta) - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
We have \(f(p-u) = f(p-v)  \leqslant k\) and \(p-u + p-v > p\), the extra condition on \(f\) provides \(f(p-u-v) \neq f(p-u)\).
We assumed \(h(x+y) \leqslant k\) hence \(h(x+y) = f(p-u-v)\). Since \(h(x) = f(p-u)\), we have at last \(h(x+y) \neq h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1) - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
We assumed \(h(x+y) \leqslant k \) hence \(h(x+y) = f( p- (p-u-v)) = f((p-u + p-v))\). Since \(f(p-u+p-v) \neq f(p-u)\),
we get \(h(x+y) \neq h(x)\).
\end{itemize}

We now assume \(h(x) > k\) (and also \(h(x+y) > k\) otherwise the expected result is trivial). 
Then \(h(x) = k + g(\alpha) = k + g(\beta) = h(y)\), hence \(g(alpha) = g(\beta\). We have the two same cases as before.

\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta) - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
Since \(g(alpha) = g(\beta)\), the sum-free condition garanties \(g\alpha) \neq g(\alpha+\beta)\), 
thus \(h(x+y) = k + g(\alpha + \beta) \neq k + g(\alpha) = h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1) - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
Because of the assumption \(h(x) > k\), we necessarily have \(f(p-u) = k+1 = f(p-v)\). 
Then \(f(p-(u+v-p)) = f(p-u + p-v) \leqslant k \) with the sum-free condition on \(f\), 
hence \(h(x+y) = f(p-(u+v-p)) \) by construction of \(h\). Thus \(h(x+y) \leqslant k < h(x)\).
\end{itemize}

### 2021-04-08 article/weakschur.tex

In this section, we generalize Rowley's constructions in [2]. We then introduce, by analogy with the third section, the
integer \(WS^+(n)\)
to build suitable templates.

@@

Up to now, there was no equivalent for weak Schur numbers of Abott and Hanson's construction [refenrece]. Here we
give a general lower bound for weak Schur numbers as a function of both regular and weak Schur numbers.
The following theorem, inspired by Rowley's inequalities for \(WS(n+1)\) and \(WS(n+2)\), was found and proved
by Romain Ageron.

@@

\begin{remark}
This formula includes the results of Rowley \hyperlink{label2}{\textbf{[2]}} as a special case. For \(n>2\),
this formula does not give new lower bounds but in the same way as we introduced \(S^+\)\hyperref[SE]{(Definition 3.1)},we will define \(WS^+\) and find inequalities between \(WS^+\),\(WS\) and \(S\)
\end{remark}

@@

With \( S(9) \geqslant 17803 \), we found a new lower bound for
\(WS(10)\) using Rowley's inequality. Moreover, the third inequality
gives new lower bounds for \(WS(9)\) and \(WS(14)\). However, the last
inequality doesn't give any better lower bound, even beyond \( n = 15 \)
: the best bounds are always provided by the first three. We highly suspect that these values can be improved by
investigating the search
space further, which would provide new, more effective templates. One
may try to go over this search space using a Monte-Carlo method, as in (ref Bouzy),
but with a different search space (as we explain in \hyperref[SAT]{SAT section}).
This could be the suject of a future work.

>>>>

In this section, we generalize Rowley's constructions in \cite{RowleyWS}. We then introduce, by analogy with the third section, the
integer \(WS^+(n)\)
to build suitable templates.

@@

Up to now, there was no equivalent for weak Schur numbers of Abott and Hanson's construction \cite{AbbottHanson}. Here we
give a general lower bound for weak Schur numbers as a function of both regular and weak Schur numbers.
The following theorem, inspired by Rowley's inequalities for \(WS(n+1)\) and \(WS(n+2)\), was found and proved
by Romain Ageron.

@@

\begin{remark}
This formula includes the results of Rowley \cite{RowleyWS} as a special case. For \(n>2\),
this formula does not give new lower bounds but in the same way as we introduced \(S^+\)\hyperref[SE]{(Definition 3.1)},we will define \(WS^+\) and find inequalities between \(WS^+\),\(WS\) and \(S\)
\end{remark}

@@

With \( S(9) \geqslant 17803 \), we found a new lower bound for
\(WS(10)\) using Rowley's inequality. Moreover, the third inequality
gives new lower bounds for \(WS(9)\) and \(WS(14)\). However, the last
inequality doesn't give any better lower bound, even beyond \( n = 15 \)
: the best bounds are always provided by the first three. We highly suspect that these values can be improved by
investigating the search
space further, which would provide new, more effective templates. One
may try to go over this search space using a Monte-Carlo method, as in \cite{Bouzy2015AnAP},
but with a different search space (as we explain in \hyperref[SAT]{SAT section}).
This could be the suject of a future work.
